---
title: "R Notebook"
output:
  html_document:
    df_print: paged
---

illustrate summation, subtraction, multiplication and division operations on vectors

```{r}
# Define two vectors
a <- c(10, 20, 30, 40)
b <- c(2, 4, 6, 8)

# Summation
sum_vec <- a + b
print(paste("Summation: ", toString(sum_vec)))

# Subtraction
sub_vec <- a - b
print(paste("Subtraction: ", sub_vec))
cat("sub",sub_vec)

# Multiplication
mul_vec <- a * b
print(paste("Multiplication: ", toString(mul_vec)))

# Division
div_vec <- a / b
print(paste("Division: ", toString(div_vec)))

```

```{r}
# Create two matrices
matrix1 <- matrix(c(1, 2, 3, 4, 5, 6), nrow = 2, ncol = 3, byrow = TRUE)
matrix2 <- matrix(c(7, 8, 9, 10, 11, 12), nrow = 3, ncol = 2, byrow = TRUE)

# Display the matrices
cat("Matrix 1:\n")
print(matrix1)

cat("\nMatrix 2:\n")
print(matrix2)

# Perform matrix multiplication
result <- matrix1 %*% matrix2

# Display the result
cat("\nResult of Matrix Multiplication:\n")
print(result)

```

### 1. **Matrix--Vector Multiplication**

-   A matrix $A$ of size m×n can be multiplied with a vector $x$ of size $n\times 1$.

-   The result is a new vector $y$ of size $m\times 1$.

-   Formula:

$y=A⋅x$

where

$$
y_i = \sum_{j=1}^n A_{ij} \cdot x_j
$$​

```{r}
A <- matrix(c(1,2,3,4,5,6), nrow=2, byrow=TRUE)
x <- c(1,2,3)
print(A)
print(x)
y <- A %*% x
print(y)

```

### **Matrix--Vector Element-wise Multiplication**

-   If vector length matches matrix dimension, element-wise multiplication occurs.

-   Formula:

$$C_{ij}=A_{ij}⋅x_j$$

```{r}
A <- matrix(c(1,2,3,4), nrow=2, byrow=TRUE)
x <- c(10,20)

result <- A * x   # vector recycled across rows
print(result)

```

```{r}
# Create some sample data
data <- c(10, 12, 15, 18, 20, 22, 25, 28, 30, 32, 35)

# Create a simple box plot
boxplot(data)

# Create a box plot with a title and axis labels
boxplot(data, main = "My Box Plot", xlab = "Data Points", ylab = "Values")

# Create box plots for different groups using a formula
# Example: comparing 'value' based on 'group'
 data_frame <- data.frame(value = c(1:10, 15:24), group = rep(c("A", "B"), each = 10))
 boxplot(value ~ group, data = data_frame, main = "Box Plot by Group")
```

```{r}
#data <- rnorm(100)
qqnorm(data)

```

```{r}
hist(data,breaks=3)
hist(data, breaks=c(0,5,10,15,20,25,30,35))
```

```{r}
# Sample data
x <- c(1, 2, 3, 4, 5)
y <- c(2, 4, 6, 8, 10)

# Create a scatter plot
plot(x, y, 
     main = "Scatter Plot in Base R",  # Title of the plot
     xlab = "X-axis Label",            # Label for the X-axis
     ylab = "Y-axis Label",            # Label for the Y-axis
     col = "blue",                     # Color of the points
     pch = 2)                          # Shape of the points

```
#subsetting

```{r}
mixed<-c(1,'a')
mixed
```



```{r}
remain<-c(11,12,11,13)
 remain
 
 suits<-c("Spades", "hearts", "diamonds", "clubs")
names(remain) <-suits
remain

```

```{r}
remain1<-c(Spades=11,hearts=12,diamonds=11,clubs=13)
remain1

```
```{r}
remain[1]

remain['Spades']

```

```{r}
remain2<-remain[c(1,4)]
remain2

```

```{r}
remain[c("clubs", "spades")]







```

```{r}
remain[c(FALSE,TRUE,FALSE,TRUE)]

```


```{r}  
remain[-1]

```

```{r}
sel_vec<-remain==11
print(sel_vec)
remain[sel_vec]



```



```{r}
remain[c(TRUE,FALSE)]
```
#Matrix

```{r}
m=matrix(1:6,nrow=2)
m
```


```{r}

m[,3]

m[1,]

m[1,1]


```


```{r}
m
m[2,c(2,3)]

```

```{r}
a <- c(10, 20, 30, 40,50)
summary(a)
```

