---
title: "CONDITIONALS, CONTROL FLOW and ITERATIVE"
output:
  html_document:
    df_print: paged
---

Sum of numbers

```{r}
# Install the package if you haven't already
#install.packages("remotes")
#remotes::install_github("mkearney/rmd2jupyter")
#library(rmd2jupyter)
#rmd2jupyter("minor.Rmd") #,output_file="m.ipynb")
sum=0
a=c(1,3,5,10,7)
for( num in a){
  sum=sum+num
}
print(sum)
```

Maximum of numbers

```{r}
max=a[1]
print(a)
for (num in a){
 # print(num)
  if ( max <= num ) {
    max=num
  }
}
print(max)
print(max(a))
```

```{r}
# List of strike rates
b <- list(100, 123, 34, 99)
good<-c()
bad<-c()
if(b[1]>100){
  good<-c(good,1)
} else{
  bad<-c(bad,1)
}


if(b[2]>100){
  good<-c(good,2)
} else{
  bad<-c(bad,2)
}

i<-4
if(b[i]>100){
  good<-c(good,i)
} else{
  bad<-c(bad,i)
}

# Convert list to numeric vector for easier processing
strike_rates <- unlist(b)

# Categorize using ifelse
categories <- ifelse(strike_rates > 100, "good", "bad")

# Display results
data.frame(StrikeRate = strike_rates, Category = categories)

```

# 4. CONDITIONALS AND CONTROL FLOW

## Relational operators

```{r}
TRUE==T
#TRUE=TRUE  results in error
TRUE==TRUE
```

```{r}
'HELLO'=="GoodBye"
#3===2
TRUE!=T
TRUE!=F
"hello"!="GoodBye"
```

```{r}
3<5
3>5
"Hello">"Goodbye"
T<F
```

```{r}
"Hello"<"hello"
5>=3
3>=3
```

## Relational operators and vectors

```{r}
linkedin<-c(16,9,13,5,2,17,14)
linkedin>10
facebook<-c(17,7,5,16,8,13,14)
facebook

facebook<=linkedin
```

## Logical operators

### AND operator

```{r}
TRUE & TRUE
TRUE & FALSE
FALSE & TRUE
FALSE & FALSE

```

```{r}
# 
x<-12
x>5 & x<15

x<-17
x>5 & x<15
```

### OR operator

```{r}
TRUE | TRUE
TRUE | FALSE
FALSE | TRUE
FALSE | FALSE

```

```{r}
y<-4
y<5 | y>15
y<-14
y<5 | y>15
```

### NOT operator

```{r}
!TRUE

!FALSE

!(x<5)
```

```{r}
is.numeric(5)
!is.numeric(5)
is.numeric("hello")
!is.numeric("hello")

```

## Logical operators and vectors

```{r}
a=c(TRUE,TRUE,FALSE)
b=c(TRUE,FALSE,FALSE)
a&b

a|b

!a
```

## Conditional Statements

if(condition){ #\<\<\<\< note the placement of brackets expression } -----------------

if(condition){ expression 1 } else { #\<\<\<\< note the placement of brackets expression 2 }

```{r}
#find whether the given number is negative or not positive.

x<-5

if(x<0){
  print("x is a negative number")
} else {
  print("x is either a positive number or zero")
}


```

# 5. Iterative programming in R

Most of the problems that we come across, iterative statements are necessary.

For and while loops

## WHILE loop

while(test_expression){ statement }

```{r}
v<-"Hello"
cnt<-2
while(cnt<7){
  print(v)
  cnt<-cnt+1
}
```

## FOR loop

syntax

for ( variable in start:end){ statement }

```{r}
for (i in 2:7){
 # print(v)
  cat(v,i, "\n")
}
print(1:5)
```

## looping over a list & vector

```{r}
#vector
v<-c(1,2,"One", "True")
for (i in v) 
  { print(i)}

```

```{r}
a<-1
b<-1
c=a+b
print(c)
for (i in 1:20){
  
  a<-b+c
  b<-c+a
  c<-a+b
  
  print(a)
  print(b)
  print(c)
}
```



```{r}
v<-c(1,2,"One", "True")
cnt<-1
while( cnt<length(v))
{
  print(v)
  cnt<-cnt+1
}
```

### loops for matrices

```{r}
y<-matrix(1:20, nrow=5,ncol=4)
for( i in 1:nrow(y)){
  for(j in 1:ncol(y)){
    #print(y[i,j])
    #y[i][j]
    cat(y[i,j], "\n ")
  }
  cat("\n")
}
print(y)
cat(y)
```

### loops for data frames

```{r}
patientID<-c(1,2,3,4)
age<-c(25,34,28,52)
diabetes<-c("Type1","Type2", "Type1", "Type1")
status<-c("Poor","Improved", "Excellent", "Poor")
patientdata<-data.frame(patientID,age,diabetes,status)
for (i in names(patientdata)){
  print(i)
}

for (i in names(patientdata)){
  print(patientdata[i])
}
```

###Loops for lists

```{r}
new<-list(one=c("11","11","11"), two=c("22","22","22"))
for( i in names(new))
{
  print(new[i])
}
```

```{r}
l1<-list(a=1) #,2,3,4,5)
for( i in l1){
  print(i)
}

for( i in names(l1)){
  print(i)
  print(l1[i])
}
```

| Function        | Type      | Default Values | Example Use            |
|-----------------|-----------|----------------|------------------------|
| `numeric(10)`   | double    | `0`            | decimals, real numbers |
| `integer(10)`   | integer   | `0L`           | whole numbers, IDs     |
| `character(10)` | character | `""`           | text, labels           |

```{r}
f<-numeric(10)
print(f)
f[12]<-2
print(class(f))
print(f)
```

```{r}
# Write a R program to print the Fibonacci series
len<-10
#fibvals<-numeric(len)
#fibvals[1]<-1
fibvals<-numeric(len)
fibvals[1]<-0
fibvals[2]<-1
for( i in 3:len) {
  fibvals[i]<-fibvals[i-1]+fibvals[i-2]
}
fibvals
```

```{r}
# write a R program to print the first 100 prime numbers
is.prime1<-function(num){
  if(num==2){
    return (TRUE)
  } 
  
  l1=2:(num-1)
  cat("l1=",l1)
  t=num%%l1
  cat("\n t")
  print(t)
  cat("\n (t==0)")
  print(t==0)
  if( any(t==0)){
    return(FALSE)
  } else{
    return(TRUE)
  }
}
a=is.prime1(31)
print(a)
```

```{r}
is_prime <- function(num) {
  if (num <= 1) return(FALSE)
  if (num == 2) return(TRUE)
  !any(num %% 2:(num-1) == 0)
}

# Test
is_prime(7)   # TRUE
is_prime(9)   # FALSE
is_prime(2)   # TRUE
is_prime(1)   # FALSE

```

```{r}
num<-12
flag<-0
for(i in 2:(num-1)){
  if(num%%i==0){
    flag<-2
  }
}
print(i)
if(flag==0){
print("Prime")
} else{
  print("not prime")
}
```

Sum of numbers
```{r}

n<-5:10
sum<-0
for (i in n){
  sum<-sum+i
}
print(sum)
cat("size of n is ", length(n), "\n")
sum<-0
for ( i in 1:6){
  sum<-sum+n[i]
}
print(sum)
```

find the GCD
```{r}
a=18
b=20
for(i in 1:20){
  a=b%%a
  b=a%%b
  cat("a=",a,"\n")
  cat("b=",b,"\n")
  
}
```
```{r}


a=18
b=20
while (b != 0) {
        temp <- b;
        b <- a %% b;
        a <- temp;
    }
print(a)
```

```{r}
x <- c("Low", "Medium", "High", "Medium", "Low")
just_factor<-factor(x)
ordered_x <- factor(x, levels = c("Low", "Medium", "High"), ordered = TRUE)
## TRY the above with FALSE also.
print("ordered x is")
print(ordered_x)
print(as.integer(ordered_x))
print("just factor is ")
print(just_factor)
print(as.integer(just_factor))
print("-----------------------------------------")
print(as.integer(ordered_x))
print(max(ordered_x))

print(max(ordered_x))

```

