---
title: "Functions"
output:
  html_document:
    df_print: paged
---

function_name \<- function(arguments){ computations on the arguments task-specific code statements } \### User defined functions

```{r}
#Function Definition
func_square<-function (x){
  y<-x*x
  print(y)
}
# function call
func_square (5)
```

```{r}
#Function Definition
func_square<-function (x){
  x*x
  #print(y)
}
# function call
y<-func_square (5)
y
```

```{r}
func_square<-function (x){
  return (x*x)
  #print(y)
}
# function call
y<-func_square (5)
y
```

```{r}
f<-function(x){
  g<-function(x){
    #x<-10
    print("inside g")
    print(x)
  }
  g(5)
  print("Inside of F")
  print(x)
}
f(10)

```

```{r}
a<-10
outer_func<-function(){
  inner_func<-function(){
    a<<-30
    print(a)
  }
  inner_func()
  print(a)
  
}
outer_func()
print(a)
```

```{r}
multi<-function(x,y=NULL){
  if(!is.null(y)) x<-x*y
  return(x)
}
multi(2)
multi(2,3)
```

### Complex numbers

```{r}
c<-complex(real=2,imaginary = 5)
f<-complex(real=4, imaginary = 7)
g<-c+f
print(g)
is.complex(g)
```

```{r}
c*f
```

### Quick Sort

```{r}
qsort<-function(value){
  # computer till value =1
  if(length(value)<=1){
    return(value)
  }
  data<-value[1]
  div<-value[-1]
  
  # re-arrange the values i.e less values than data values come before,
  #and all greater values come after
  v1<-div[div<data]
  v2<-div[div>=data]
  
  v1<-qsort(v1) # do quick sort for v1 vaues
  v2<-qsort(v2)
  
  return(c(v1,data,v2))
  
}

qsort(c(10,3,44,77))
```

```{r}
ins <- function (hdidx, tree, newval, inc){
tr <- tree

# check for room to add a new element
tr$nxt <- tr$nxt + 1
if (tr$nxt > nrow(tr$mat))
tr$mat <- rbind(tr$mat, matrix(rep (NA, inc*3),nrow=inc,ncol=3))
newidx <- tr$nxt
# where we'11 put the new tree node
tr$mat [newidx, 3] <- newval
idx <- hdidx # marks our current place in the tree
node <- tr$mat [idx,]
nodeval <- node [3]
while (TRUE) {
# which direction to descend, left or right?
if (newval <= nodeval) dir <-1 else dir <- 2
# descend
# null link?
if (is.na(node [dir])) {
tr$mat [idx, dir] <- newidx
break
}else {
idx <- node [dir]
node <- tr$mat [idx,]
nodeval <- node [3]
}
}
return(tr)
}
```

### Recursion

```{r}
fact<-function(n){
  f<-1
  for ( i in 1:n){
    f=f*i
  }
  return(f)
}
result<-fact(6)
print(result)
```

```{r}
rec_fact<-function(x){
  if(x==0) {
    return(1)
  }else{
    return(x*rec_fact(x-1))
  }
}
rec_fact(6)
```

# mathematical functions in R

```{r}
abs(8/-22)
abs(-4)
```

log function returns the natural logarithm (to base e) of a numeric value

```{r}
log(5)
log(4)
```

```{r}
log(10,10) # for base 10
log(100,10)
```

exp function --- e raised to the power of numeric value

```{r}
exp(1.609)

```

log10 returns the logarithm of numeric value with base 10

```{r}
log10(5)
```

```{r}
sqrt(64)
```

```{r}
factorial(7)
```

```{r}
round(46.26,0)
round(4.26,1)
round(9.3443884,4)
```

signif(numeric, digit)-- gives the first digit significcant integers of the numeric value

```{r}
signif(9.343434,4)
signif(pi,3)
```

In R, `runif()` and `rnorm()` are functions used to generate random numbers from specific probability distributions.

-   `runif()` (Uniform Distribution):

    -   This function generates random numbers that are uniformly distributed within a specified range. In a uniform distribution, every value within the defined interval has an equal probability of being selected.

    -   **Syntax:** `runif(n, min = 0, max = 1)`

        -   `n`: The number of random values to generate.

        -   `min`: The lower limit of the uniform distribution (default is 0).

        -   `max`: The upper limit of the uniform distribution (default is 1).

```{r}
runif(5, min = 10, max = 20)
runif(5)
```

`rnorm()` (Normal Distribution):

-   This function generates random numbers that follow a normal (Gaussian) distribution. In
    a normal distribution, values cluster around the mean, with the
    probability of observing a value decreasing as it moves further from the
    mean, according to the standard deviation.

-   **Syntax:** `rnorm(n, mean = 0, sd = 1)`

    -   `n`: The number of random values to generate.

    -   `mean`: The mean (average) of the normal distribution (default is 0).

    -   `sd`: The standard deviation of the normal distribution (default is 1).

```{r}
rnorm(5, mean = 50, sd = 10)
rnorm(5)
```
